---
title: "dc_residential_anova"
author: "Mojahid Osman"
date: "March 1, 2019"
output: html_document
---


```{r pressure, echo=FALSE}

library(dplyr)
library(ggplot2)

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
#The Question we need to answer is: Is there any price different between houses prices in different Quadrants. 

dc.data <- read.csv("~/Documents/Personal/Master/Intro_Data_Science/Project/Data/dc_residential_data_clean.csv", header=T)
View(dc.data)

#The Null Hypothese:
#The mean price for all quadrant are the same : m1=m2=m3=m4
#Alternative: At lease one mean price is different from the other 

```{r}
ggplot(dc.data, aes(QUADRANT,PRICE) )

plot(dc.data$PRICE~dc.data$QUADRANT)

table(dc.data$QUADRANT)

table(price_outlier$QUADRANT)

#Count of Houses by Quard
table(dc.data$QUADRANT)

plot(dc.data$PRICE~dc.data$QUADRANT)

```

#Get the mean price for each quadrant to get some idea about house prices
aggregate(PRICE~QUADRANT,dc.data,mean)
aggregate(PRICE~QUADRANT,dc.data,sd)
#Lets do ANOVA test now  
model <- aov(dc.data$PRICE~price.quadrant$QUADRANT)
summary(model)
#P value is very small so that  will reject the Null hyposises, that mean the prices is signifacntly diffrent 

#price.quadrant$QUADRANT     3 3.784e+14 1.261e+14   623.8 <2e-16 ***
#Residuals               21379 4.323e+15 2.022e+11                   
#Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

# We need to know wich quardant prices is different ?
# We nned to run a post-hoc analysis, will use Tukey post-hoc .
TukeyHSD(model)
#All the P values from the result of Tukey analysis are very small which  







































