---
title: "dc_residential_anova"
author: "Mojahid Osman"
date: "March 1, 2019"
output: html_document
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r basicfcn, include=F}
# can add quietly=T option to the require() function
loadPkg = function(x) { if (!require(x,character.only=T, quietly =T)) { install.packages(x,dep=T,repos="http://cran.us.r-project.org"); if(!require(x,character.only=T)) stop("Package not found") } }
```

```{r pressure, echo=FALSE}

library(dplyr)
library(ggplot2)

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
#The Question we need to answer is: Is there any price different between houses prices in different Quadrants. 

dc.data <- read.csv("~/Documents/Personal/Master/Intro_Data_Science/Project/Data/dc_residential_data_clean.csv", header=T)
View(dc.data)

#The Null Hypothese:
#The mean price for all quadrant are the same : m1=m2=m3=m4
#Alternative: At lease one mean price is different from the other 

```{r}
ggplot(dc.data, aes(QUADRANT,PRICE) )

plot(dc.data$PRICE~dc.data$QUADRANT)

table(dc.data$QUADRANT)

table(price_outlier$QUADRANT)

#Count of Houses by Quard
table(dc.data$QUADRANT)

plot(dc.data$PRICE~dc.data$QUADRANT)

```

#Get the mean price for each quadrant to get some idea about house prices
aggregate(PRICE~QUADRANT,dc.data,mean)
aggregate(PRICE~QUADRANT,dc.data,sd)
#Lets do ANOVA test now  
```{r}
model <- aov(dc.data$PRICE~dc.data$QUADRANT)
summary(model)
#P value is very small so that  will reject the Null hyposises, that mean the prices is signifacntly diffrent 

#price.quadrant$QUADRANT     3 3.784e+14 1.261e+14   623.8 <2e-16 ***
#Residuals               21379 4.323e+15 2.022e+11                   
#Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

````
# We need to know wich quardant prices is different ?

# So will run a post-hoc analysis, will use Tukey post-hoc .
```{r}
TukeyHSD(model)
#All the P values from the result of Tukey analysis are very small which  mean there no any siginficant realatetion between all mean
```

However once we get up to a significant number of variables and we want to know which most explain the variance in our dependent variable. 
```{r}
loadPkg("leaps")
loadPkg("ISLR")
#This is essentially best fit 
reg.dc.data1 <- regsubsets(dc.data$PRICE~., data = dc.data, nvmax = 12, nbest = 1, method = "seqrep")  
# leaps, regsubsets: Model selection by exhaustive search, forward or backward stepwise, or sequential replacement
#The plot will show the Adjust R^2 when using the variables across the bottom
plot(reg.dc.data1, scale = "adjr2", main = "Adjusted R^2")
# In the "leaps" package, we can use scale=c("bic","Cp","adjr2","r2")
plot(reg.dc.data, scale = "bic", main = "BIC")
plot(reg.dc.data, scale = "Cp", main = "Cp")
summary(reg.best)
```

#Creat a Line Chart to show timeserise of the PRICES over the time colored by Quardnet 
```{r}
price.all<- subset (select(dc.data, PRICE,QUADRANT,SALEDATE))
View(price.all)

# convert factor to numeric for convenience 
price.all$SALEDATE <- as.numeric(price.all$SALEDATE ) 
maxprice <- max(price.all$SALEDATE )


# get the range for the x and y axis 
xrange <- range(price.all$SALEDATE) 
xrange
yrange <- range(price.all$PRICE) 
yrange


# set up the plot 
plot(xrange, yrange, type="n", xlab="Sale Date", ylab="Price" ) 
colors <- rainbow(maxprice) 
linetype <- c(1:maxprice) 
plotchar <- seq(18,18+maxprice,1)


# add lines 
for (i in 1:maxprice) { 
  price11 <- subset(price.all, SALEDATE==i) 
  lines(price.all$PRICE,price.all$SALEDATE, type="b", lwd=1.5,
    lty=linetype[i], col=colors[i], pch=plotchar[i]) 
} 






price.ne<- subset (select(dc.data, PRICE,QUADRANT,SALEDATE), dc.data$QUADRANT=="NE")
View(price.ne)
price.nw<- subset (select(dc.data, PRICE,QUADRANT,SALEDATE), dc.data$QUADRANT=="NW")
View(price.nw)
price.se<- subset (select(dc.data, PRICE,QUADRANT,SALEDATE), dc.data$QUADRANT=="SE")
View(price.we)
price.sw<- subset (select(dc.data, PRICE,QUADRANT,SALEDATE), dc.data$QUADRANT=="SW")
View(price.sw)

plot(price.ne$PRICE~price.ne$SALEDATE,type="l",col="red",xlab = "DATE" ,ylab = "PRICE")

```

#Count of houses by Quardnet.
```{r}
QUADRANT<-as.factor(dc.data$QUADRANT)
count.house<-length(dc.data$PRICE)
count.house

count.house

house.counts.q <- table(dc.data$QUADRANT)
house.counts.q

barplot(house.counts.q, main="House Distribution", xlab="QUADRANT", col=c("lightblue"))

pie(house.counts.q,  main="Pie Chart of House Distribution")


```

#Some analysis based on landarea 
#Create new variables for area category (Small, Medium , Big)
```{r}
#get idea baout the area mean
summary(dc.data$LANDAREA)
aggregate(LANDAREA~QUADRANT,dc.data,mean)

dc.data$HouseSize[dc.data$LANDAREA<=6]<- 'Small'
dc.data$HouseSize[dc.data$LANDAREA>6 & dc.data$LANDAREA<=8]<- 'Medium'
dc.data$HouseSize[dc.data$LANDAREA>8]<- 'Big'

house.counts.q.housesize <- table(dc.data$HouseSize,dc.data$QUADRANT)
house.counts.q.housesize

barplot(house.counts.q.housesize, main="House Distribution by LAND AREA",
  xlab="QUADRANT", col=c("lightblue","lightyellow","red"), beside=TRUE,legend = rownames(house.counts.q.housesize))

```
#Some analysis based on Price 
#Create new variables for Price Group  (Low, Medium , High)

#get idea baout the area mean
```{r}
summary(dc.data$PRICE)
aggregate(PRICE~QUADRANT,dc.data,mean)

dc.data$Pricegroup[dc.data$PRICE<=400000]<- 'Low'
dc.data$Pricegroup[dc.data$PRICE>400000 & dc.data$LANDAREA<=600000]<- 'Medium'
dc.data$Pricegroup[dc.data$PRICE>600000]<- 'High'

house.Pricegroup <- table(dc.data$Pricegroup,dc.data$QUADRANT)
house.Pricegroup 


barplot(house.Pricegroup, main="House Distribution by Price",
  xlab="QUADRANT", col=c("lightblue","lightyellow","red"), beside=TRUE,legend = rownames(house.Pricegroup))



```






























