---
title: "DATS6101 HW 3"
author: "Chris Montgomery"
date: "2/1/2019"
output: html_document
---




```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## DATS 6101 Homework III

```{r include = F}
library(dplyr); library(MASS)

df <- Pima.te

```

#Q1.

```{r}
summary(df)

```

#Q2.

```{r}
str(df)
```

#Q3.

```{r}
colnames(df)
```

#Q4 and Q5.

```{r}
df[,"age"] <- as.numeric(df[,"age"])
summary(df[c("bmi", "age")])
length(na.omit(df$age)); length(na.omit(df$bmi))

range(df$bmi); range(df$age)

nrow(df)

```
There are a total of 332 observations (women) in the dataset. 

#Q6.

```{r}
df[1:5,1:4]


```

#Q7.

```{r}
df[df$bmi >= 50,]
df %>% filter(bmi >= 50)
```

#Q8.

```{r}
nrow(df[df$type == "Yes",]) / nrow(df)
```

The 32.8% of individuals within the sample data have diabetes as defined by WHO criteria. 
#Q9.

```{r}
hist(df$bmi)
```

#Q10. 
```{r}
t.test (df$bmi, conf.level = .8) ; t.test(df$bmi, conf.level = .99) 
```

Assuming valid T-distribution assumptions, we are 80% confident that the mean BMI for our population is between 32.7 and 33.8. Likewise, we are 99% confident the population mean falls between 32.2 and 34.3. 

 
#Part II

#Q11.
```{r}
df <- read.csv("hw3/vlbw.csv")

```

#Q12
```{r}
hist(df$hospstay)


#Let's remove observations using Tukey's 1.5 IQR method. 
outliers <- function(data.variable){

data.correct <- na.omit(data.variable[data.variable > -1])

iqr <- quantile(data.correct, 0.75, na.rm = TRUE) - quantile(data.correct, 0.25, na.rm = TRUE) 

lower.bound <- quantile(data.correct, .25, na.rm = TRUE) - (1.5 * iqr); lower.bound
upper.bound <- quantile(data.correct, .75, na.rm = TRUE) + (1.5 * iqr); upper.bound 

data.correct <- data.correct[!data.correct > upper.bound & !data.correct < lower.bound]


return(hist(data.correct, main = "Histogram of Outlier Corrected Data"))
return(data.correct)}; outliers(df$hospstay) 
```

#Q13 and #Q14

```{r}
boxplot(df$lowph); hist(df$lowph)
```

#Q15.

```{r}
 t.test(df$lowph, conf.level = .99)
```


