---
title: "EDA, z-score, variance"
author: "GWU Intro to Data Science DATS 6101"
date: "Jan 30, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r basicfcn, include=F}
# can add quietly=T option to the require() function
loadPkg = function(x) { if (!require(x,character.only=T, quietly =T)) { install.packages(x,dep=T,repos="http://cran.us.r-project.org"); if(!require(x,character.only=T)) stop("Package not found") } }
```


## EDA

Let's load our data and create a dataframe call **mlb**, then take a look.

```{r, include=F}
getwd()
# mlb <- data.frame(read.csv("BaseballHeightWeight.csv", header = TRUE))
mlb <- read.csv("BaseballHeightWeight.csv", header = TRUE)
str(mlb)
head(mlb)
```

Ok let's first rename a couple of columns. You can use "names" or "colnames" to change column names like the codes here. Also notice that "=" and "<-" are mostly the same as well. Next, do some EDA.


```{r, echo=F}
colnames(mlb)[4:6]=c("height","weight","age")
# The line above produces no output. To see the changes, use str() or head()

#Start with individual variables of Height Weight and Age
head(mlb$height)
hist(mlb$height)
hist(mlb$weight)
shapiro.test(mlb$height) # test for normality
boxplot(mlb$height)
qqnorm(mlb$height) 
qqline(mlb$height)
sd.height <- sd(mlb$height)
mean.height <- mean(mlb$height)
# sd.weight <- sd(mlb$weight)   # error if there are N/A values
sd.weight <- sd(mlb$weight, na.rm = TRUE)
mean.weight <- mean(mlb$weight, na.rm = TRUE)
mean.weight
sd.weight
sd.height
```

Handy Outlier Function

```{r outlierKD, echo=F}
outlierKD <- function(dt, var) { 
     var_name <- eval(substitute(var),eval(dt))
     na1 <- sum(is.na(var_name))
     m1 <- mean(var_name, na.rm = T)
     par(mfrow=c(2, 2), oma=c(0,0,3,0))
     boxplot(var_name, main="With outliers")
     hist(var_name, main="With outliers", xlab=NA, ylab=NA)
     outlier <- boxplot.stats(var_name)$out
     mo <- mean(outlier)
     var_name <- ifelse(var_name %in% outlier, NA, var_name)
     boxplot(var_name, main="Without outliers")
     hist(var_name, main="Without outliers", xlab=NA, ylab=NA)
     title("Outlier Check", outer=TRUE)
     na2 <- sum(is.na(var_name))
     cat("Outliers identified:", na2 - na1, "n")
     cat("Propotion (%) of outliers:", round((na2 - na1) / sum(!is.na(var_name))*100, 1), "n")
     cat("Mean of the outliers:", round(mo, 2), "n")
     m2 <- mean(var_name, na.rm = T)
     cat("Mean without removing outliers:", round(m1, 2), "n")
     cat("Mean if we remove outliers:", round(m2, 2), "n")
     response <- readline(prompt="Do you want to remove outliers and to replace with NA? [yes/no]: ")
     if(response == "y" | response == "yes"){
          dt[as.character(substitute(var))] <- invisible(var_name)
          assign(as.character(as.list(match.call())$dt), dt, envir = .GlobalEnv)
          cat("Outliers successfully removed", "n")
          return(invisible(dt))
     } else{
          cat("Nothing changed", "n")
          return(invisible(var_name))
     }
}
```


Running the outlierKD code below triggers a prompt on the console window (yes/no). Answer there to continue.
```{r}
outlierKD(mlb, height)
```


We can also run the vairance
```{r, echo=TRUE, echo=F}
var(mlb$height, na.rm = T)
mean(mlb$height, na.rm = T)
sd(mlb$height, na.rm = T)
```
Now that we know the mean and std we can generate a z score for individual heights and probabilities (assuming normal distribution)
```{r, echo=F}
z80height <- (80-mean.height)/sd.height
z80height
```

Note that now we have a z score, we could look it up on a table or use R to calculate the percentage (assumed normal dist), pnorm tells us the percentage.
```{r, echo = F}
pnorm(z80height)
pnorm(2.73)
```

Better yet, we can convert raw scores into z-scores easily for all the data:
```{r, echo=TRUE}
mlb$zheight = scale(mlb$height)
mlb$zweight = scale(mlb$weight)
```

Notice that the dataframe mlb has two more columns/variables now. 

Let us simply check that zheight and zweight are centered at zero, with sd = 1
```{r, echo=F}
mean(mlb$zheight, na.rm = TRUE)
sd(mlb$zheight, na.rm = TRUE)

summary(mlb$zheight)
summary(mlb$zweight)
```

And try this describe function:
```{r, echo = F, message=FALSE, warning=FALSE}
loadPkg("Hmisc") # this package contains the describe function below
describe(mlb$weight)
str(mlb$weight)
str(mlb$zweight)
```

It's a nicer way to describe the data, then say, summary( ). But describe( ) does not work on the entire dataframe. Only one vector at a time.

## Confidence Interval / Z-Interval / T-Interval
Now let's try these inferential statistics tools that we learned:  
z-interval  
t-interval  

```{r, include=F}
mean.height = mean(mlb$height, na.rm = TRUE); 
mean.weight = mean(mlb$weight, na.rm = TRUE); 
```

Before we do that, let us record again the mean height and weight for the entire mlb population. The mean height is mean.height = `r format(mean.height, digits=5)` inches, mean.weight = `r format(mean.weight, digits=5)` lbs. 

Now let us create a smaller sample of the mlb players out of the entire population, and call it mlbsample.
```{r mlbsample, include = F}
set.seed(123) # just so that everyone have the same sample for comparison
mlbsample = mlb[ sample(nrow(mlb),30), ]
str(mlbsample)
head(mlbsample)
format(mean(mlbsample$height), digits=5)
format(mean(mlbsample$weight, na.rm = TRUE), digits=5)
```

### Z-Interval Estimation (and Z-Test for next class)
```{r z_test1}
loadPkg("BSDA") # for z.test
ztest95 = z.test(x=mlbsample$height, sigma.x = 2.31) # default conf.level = 0.95
ztest95
ztest99 = z.test(x=mlbsample$height, sigma.x = 2.31, conf.level=0.99 )
ztest99
ztest50 = z.test(x=mlbsample$height, sigma.x = 2.31, conf.level=0.50 )
ztest50
```

Let us also take a look at the object "ztest99". It contains these info:
```{r z_test2}
names(ztest99)
ztest99$conf.int
# ztest99$alternative
ztest99$estimate
ztest99$statistic
ztest99$method
# etc etc
```

Now let us compare the three intervals, and note the actual average for the entire population is `r mean.height`.
```{r z_test2b}
mean.height
ztest99$conf.int
ztest95$conf.int
ztest50$conf.int
```

## T-Intervals Estimation (and T-Test for next class)
Similar for t-test, except it is **easier**, since we do not need to know the standard deviation (sigma) of 
the population. First, the results for the default c=95% confidence level.
```{r t_test, include = F}
# t.test is included in the basic R package 'stats'
ttest95 = t.test(x=mlbsample$height) # default conf.level = 0.95
ttest95
ttest99 = t.test(x=mlbsample$height, conf.level=0.99 )
ttest99
ttest50 = t.test(x=mlbsample$height, conf.level=0.50 )
ttest50
names(ttest99)
ttest99$conf.int
ttest99$alternative
ttest99$estimate
```


Now let us compare the three intervals, with that produced from the z-test.
```{r t_test2_99}
mean.height
ztest99$conf.int
ttest99$conf.int
```

```{r t_test2_95}
mean.height
ztest95$conf.int
ttest95$conf.int
```

```{r t_test2_50}
mean.height
ttest50$conf.int
ztest50$conf.int
```
You can try the above for the weight data?


